LANGuage IDentification

### 2024-12-12 paper/paper.tex
paper: neuron

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Using Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{empheq}
\usepackage{mdframed}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{color}
\usepackage{psfrag}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{bm}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\selectlanguage{english}
\usepackage{blindtext}
\usepackage{csquotes}
\usepackage{hyperref}
\hypersetup{colorlinks,
           citecolor=black,
           filecolor=black,
           linkcolor=black,
           urlcolor=black,
           bookmarksopen=true,
           pdftex}
\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\addbibresource{references.bib}
\usepackage{bookmark}
\usepackage{enumitem}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot[C]{\copyrightnotice}
\setlength{\headheight}{14.5pt}
%\addtolength{\topmargin}{-2.5pt}
\usepackage{mathtools}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

    Let's say \hyperref[tab:rainbow_model_vocab]{our model's vocabulary} are
    the words in English for the 7 colors in the rainbow.

@@

    The way I implemented the BoW was by reading a file and collect every
    single word from it separated by white spaces, dashes, punctuation,
    parenthesis or quotation marks and then spit to \texttt{stdout} a formated
    vocabulary, so you can pipe it later or write it to a file. Words are
    case-sensitive, and it's possible to create the vocabulary in alphabetic
    order.

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Using Packages %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{geometry}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{empheq}
\usepackage{mdframed}
\usepackage{booktabs}
\usepackage{lipsum}
\usepackage{graphicx}
\usepackage{color}
\usepackage{psfrag}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{bm}
% % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % % %
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\selectlanguage{english}
\usepackage{blindtext}
\usepackage{csquotes}
\usepackage{hyperref}
\hypersetup{colorlinks,
           %citecolor=black,
           %filecolor=black,
           %linkcolor=black,
           %urlcolor=black,
           bookmarksopen=true}
\usepackage[
backend=biber,
style=alphabetic,
sorting=ynt
]{biblatex}
\addbibresource{references.bib}
\usepackage{bookmark}
\usepackage{enumitem}
\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyfoot[C]{\copyrightnotice}
\setlength{\headheight}{14.5pt}
%\addtolength{\topmargin}{-2.5pt}
\usepackage{mathtools}
\usepackage{tikz}
\usetikzlibrary{positioning,fit,calc}
\usepackage{float}
\usepackage{wrapfig}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

    \listoffigures
    %\listoftables

@@

    Let's say {our model's vocabulary}~\ref{tab:rainbow_model_vocab} are
    the words in English for the 7 colors in the rainbow.

@@

    The way I implemented the BoW was by reading a file and collect every
    single word from it separated by white spaces, dashes, punctuation,
    parenthesis or quotation marks and then spit to \texttt{stdout} a formatted
    vocabulary, so you can pipe it later or write it to a file. Words are
    case-sensitive, and it's possible to create the vocabulary in alphabetic
    order.

@@

    \section{Neuron}

    As explained at the start of chapter 3.1 of \cite{bishop}, the base for the
    simplest supervised linear regression models will look like
    \eqref{eq:linear_model}
    \begin{equation}
        \begin{split}
            y(\mathbf{x}, \mathbf{w}) & = (1)b + w_{0}x_{0} + w_{1}x_{1} +
            \ldots + w_{D}x_{D} = \\ & = (1)b +
            \displaystyle\sum_{j=0}^{D}\left(w_{j}x_{j}\right)
        \end{split}
        \label{eq:linear_model}
    \end{equation}
    where $ \mathbf{x} =
    \begin{bsmallmatrix}x_{0}&\cdots&x_{D}\end{bsmallmatrix}^{T} $, a linear
    function of the parameters $ \mathbf{w} =
    \begin{bsmallmatrix}w_{0}&\cdots&w_{D}\end{bsmallmatrix} $ and bias $b$. We
    can also make the bias part of the $ \mathbf{w} $ and at the same position
    in $ \mathbf{x} $ put the value 1 and have the equation like
    \eqref{eq:linear_model_matrix}
    \begin{equation}
        y(\mathbf{x}, \mathbf{w}) = \mathbf{x}\cdot\mathbf{w}
        \label{eq:linear_model_matrix}
    \end{equation}
    a simple matrix multiplication. Both matrixes have the same amount of
    elements $ D $.

    There is also the notion of applying what is called as \textbf{basis
    function} to the $ \mathbf{x} $ parameters before everything. So there will
    be a matrix of basis functions $ \mathbf{\phi} =
    \begin{bsmallmatrix}\phi_{0}&\cdots&\phi_{D}\end{bsmallmatrix}^{T} $ where
    $ b = x_{i} $ and $ \phi_{i}(b) = 1 $.

    If that's the case the function will look like this
    \eqref{eq:linear_model_basis_f}
    \begin{equation}
        \begin{split}
            y(\mathbf{x}, \mathbf{w}) & =
            \displaystyle\sum_{j=0}^{D}\left(w_{j}\phi_{j}(x_{j})\right) = \\
            & = \mathbf{\phi}\left(\mathbf{x}\right)\cdot\mathbf{w}
        \end{split}
        \label{eq:linear_model_basis_f}
    \end{equation}

    The functions can be seen as a way to pre-process and extract the important
    parts of the input This list can have nonlinear functions making $
    y(\mathbf{x}, \mathbf{w}) $ nonlinear too.

    In case we are working with a classification model, we might want to
    introduce an \textbf{activation function} $ f\left(\cdot\right) $ that
    after all the computations in $ y(\mathbf{x}, \mathbf{w}) $
    \eqref{eq:linear_model_basis_f} transforms that result into a value
    corresponding to being the probability of $ \mathbf{x} $ being of a certain
    class.

    And this will be what I define as an artificial neuron \eqref{eq:neuron}.
    You can see a visual representation of it on page~\pageref{fig:neuron}

    \begin{equation}
        y(\mathbf{x}, \mathbf{w}) = f \left( \mathbf{\phi} \left( \mathbf{x}
        \right) \cdot\mathbf{w} \right)
        \label{eq:neuron}
    \end{equation}

    \begin{figure}
        \begin{center}
            \begin{tikzpicture}[
                    neuron/.style={circle, draw=black!90, very thick},
                    var/.style={}, io/.style={circle, draw=black!90,
                    fill=black!10, very thick, minimum size=7mm},
                ]
                \node (neuron_logic) {$ y\left(\mathbf{x}, \mathbf{w}\right)
                $};
                \node (weights) [above=of neuron_logic] {$ \mathbf{w} $};
                \node[text opacity=0] (invisible) [below=of neuron_logic] {neuron};
                \node[neuron, fit={(weights)(neuron_logic)(invisible)}] (neuron) {};
                % \node[var] (input) [left=of neuron] {$ \mathbf{x} = \begin{bmatrix}x_{0}&x_{1}&\vdot&x_{D}\end{bmatrix} $};
                \node[var] (input) [left=of neuron] {$ \mathbf{x} $};
                \node[var] (bias) [above=of input] {$ b = x_{0} $};
                \node[var] (output) [right=of neuron] {$ y $};
                \node[io] (basis) at (neuron.west) {$
                \mathbf{\phi}\left(\mathbf{x}\right) $};
                \node[io] (activation) at (neuron.east) {$ f\left(y\right) $};

                \draw[->] (bias.south) -- (input.north);
                \draw[->] (input.east) -- (basis.west);
                \draw[->] (basis.east) -- (neuron_logic.west);
                \draw[->] (neuron_logic.east) -- (activation.west);
                \draw[->] (activation.east) -- (output.west);
                \draw[->] (weights.south) -- (neuron_logic.north);
            \end{tikzpicture}
        \end{center}
        \caption{A visual representation of a neuron.}\label{fig:neuron}
    \end{figure}

    Because I couldn't find a simple way of doing anonymous functions in my
    language of choice, \href{run:../src/utils/neuron.zig}{my implementation}
    does not allow specifying basis functions nor activation functions so every
    basis function can be thought as being the identify function $
    \mathbf{\phi}(x) = x $ and the activation function can be toggled between
    the identify and sigmoid \eqref{eq:sigmoid}.

    \begin{equation}
        \sigma\left(x\right) = \frac{1}{1+e^{-x}}
        \label{eq:sigmoid}
    \end{equation}

    %\section{Feed-Forward Neural Network}
